Brief notes on erosion

### 2016-10-16 docs/Notes.tex
Writing main routine erosion()
Got main erosion() routine working. Now working on scaline eps and sigma

We introduce the tangent angle $\theta$, defined as
\begin{equation}
\tan \theta = y_s/x_s = y_\alpha / x_\alpha
\end{equation}
Remember the identity that $x_s^2 + y_s^2 = 1$ as this is very useful.
The curvature is given by
\begin{equation}
\kappa = x_s y_{ss} - x_{ss} y_s = \pd{\theta}{s} = L^{-1} \pd{\theta}{\alpha}
\end{equation}
We are using the standard convention, where a convex shape, like a circle, has positive curvature. The normal velocity now can be written as
\begin{equation}
V_n = C \atau + \eps L^{\beta-1} \left(\pd{\theta}{\alpha} - 2 \pi \right)
\end{equation}

@@

We now discuss the numerical methods used to solve the above evolution equations. For $L$, we use an explicit method
\begin{equation}
L^{(n+1)} = L^{(n)} + \frac{\Dt}{2} \left( 3 M^{(n)} - M^{(n-1)} \right)
\end{equation}
where
\begin{equation}
M = \td{L}{t} = -\mean{\gamma}
\end{equation}
For $\theta$, we use a spectral method, giving
\begin{equation}
\td{\thh_k}{t} + 4 \pi^2 k^2 \eps L^{\beta-2} \thh_k = \hat{N}_k
\end{equation}
where the nonlinear terms are grouped together as
\begin{equation}
N = L^{-1} \left( C \pd{\atau}{\alpha} + \pd{\theta}{\alpha} V_s \right)
\end{equation}
We evolve with the integrating-factor method. When recast in terms of Gaussian filters, this gives
\begin{equation}
\thh^{(n+1)} = \GG(\theta^{(n)}, \sigma_1) 
+ \frac{\Dt}{2} \left( 3 \GG(N^{(n)}, \sigma_1) - \GG(N^{(n-1)}), \sigma_2 \right)
\end{equation}
where
\begin{align}
& \sigma_1 = 2 \pi \sqrt{\eps \Dt} \sqrt{(L^{n})^{\beta-2} + (L^{n+1})^{\beta-2} } 	\\
& \sigma_2 = 2 \pi \sqrt{\eps \Dt} \sqrt{(L^{n-1})^{\beta-2} + 2 (L^{n})^{\beta-2} + (L^{n+1})^{\beta-2} }
\end{align}

@@

\end{document}

>>>>

We introduce the tangent angle $\theta$, defined by
\begin{equation}
(x_s, y_s) = (\cos \theta, \sin \theta)
\end{equation}
so that we have
\begin{equation}
\tan \theta = y_s/x_s = y_\alpha / x_\alpha
\end{equation}
Remember the identity that $x_s^2 + y_s^2 = 1$ as this is very useful.
The curvature is given by
\begin{equation}
\kappa = x_s y_{ss} - x_{ss} y_s = \pd{\theta}{s} = L^{-1} \pd{\theta}{\alpha}
\end{equation}
We are using the standard convention, where a convex shape, like a circle, has positive curvature. The normal velocity now can be written as
\begin{equation}
V_n = C \atau + \eps L^{\beta-1} \left(\pd{\theta}{\alpha} - 2 \pi \right)
\end{equation}

@@

We now discuss the numerical methods used to solve the above evolution equations. For $L$, we use an explicit method
\begin{equation}
L^{(n+1)} = L^{(n)} + \frac{\Dt}{2} \left( 3 M^{(n)} - M^{(n-1)} \right)
\end{equation}
where
\begin{equation}
M = \td{L}{t} = -\mean{\gamma}
\end{equation}
For $\theta$, we use a spectral method, giving
\begin{equation}
\td{\thh_k}{t} + 4 \pi^2 k^2 \eps L^{\beta-2} \thh_k = \hat{N}_k
\end{equation}
where the nonlinear terms are grouped together as
\begin{equation}
N = L^{-1} \left( C \pd{\atau}{\alpha} + \pd{\theta}{\alpha} V_s \right)
\end{equation}
We evolve with the integrating-factor method. When recast in terms of Gaussian filters, this gives
\begin{equation}
\label{intfact}
\thh^{(n+1)} = \GG(\theta^{(n)}, \sigma_1) 
+ \frac{\Dt}{2} \left( 3 \GG(N^{(n)}, \sigma_1) - \GG(N^{(n-1)}), \sigma_2 \right)
\end{equation}
where
\begin{align}
& \sigma_1 = 2 \pi \sqrt{\eps \Dt} \sqrt{(L^{n})^{\beta-2} + (L^{n+1})^{\beta-2} } 	\\
& \sigma_2 = 2 \pi \sqrt{\eps \Dt} \sqrt{(L^{n-1})^{\beta-2} + 2 (L^{n})^{\beta-2} + (L^{n+1})^{\beta-2} }
\end{align}

@@

\subsection{How to scale smoothing parameters?}
We have two smoothing parameters: $\eps$ and $\sigma$. How should they scale with the simulation resolution parameters? The most important resolution parameter is $N$, the total number of points on each body, but there is also $\Delta t$ and maybe some others. Figuring out $\sigma$ is the easier part: $\sigma$ is a lengthscale and should go like $1/N$. On the other hand, $\eps$ is a diffusivity in dimensions. I did a calculation and found that $\eps$ should scale like $2/(\pi^2 N^2)$.

The calculation is based on looking at Eq.~(\ref{intfact}), where $\sigma_1 \sim 2 \pi \sqrt{\eps \Delta t}$. For the largest wavenumbers, $k_{max} \sim N/2$, we want the smoothing to beat the nonlinear term (so that overall, these modes decay). So we need $e^{-\pi^2 \eps \Delta t N^2 / 2} < 1 - C\Delta t$, which gives $\eps > 2C/(\pi^2 N^2)$, independent of $\Delta t$. If $C=1$, this gives roughly $\eps > 0.2/ N^2$.
\end{document}
